\section*{Preface}

This document is meant as an introduction to the Voss formal
verification system.
In particular, it is intended as an introduction
to using the Voss system for symbolic trajectory evaluation
and how the system can be used to support multi-level verification.
However, it is not meant to introduce the complete theory behind the system.
For this, the reader is referred to [SegBry92,SegerJoyceTwoLevel].

Since the user interface to the Voss system is a complete, fully
lazy, functional language similar to lazy-ML, this document begins
by introducing the functional language.
This is accomplished in two ways: first through a number of examples
and then as a reference guide including syntax of the language and
a description of the functions that are provided by the system. 
For someone familiar with functional programming in a
lazy language, these sections can be read very cursory.
In the second part of the manual, verification tasks using the Voss system
are explained and some secondary tools introduced.
This section is by its very nature example oriented but I have tried
to be as precise as I can.

Since the Voss system is under active development, I cannot provide any
guarantees for the correctness, suitability for any particular use.
Since I am actively developing the system, I would appreciate bug reports
and examples of dubious correctness.
In return, I'll provide as rapid fixes and updates as humanly possible.
Also, since the system is still evolving, it is quite
possible that some examples do not correspond exactly to the way
the current system works.
However, I have tried to make the manual as up-to-date as I have been able to.
In general, all the examples have been run using Voss 1.8b.

Finally, I have often been asked
what ``Voss'' really  stands for, and I have heard various attempts
containing the words Verification, Symbolic and Simulation, but
the truth is that the name was chosen specifically not to be an acronym.
Voss%
\index{Voss}%
{} is a city in Norway between Bergen and Oslo and is
prominently featured on virtually all weather forecasts in Norway.
The city is at the west end of one of the main mountain ranges in
Norway and has a special meaning to both my wife and myself.
In recognition of this, I decided to call the verification system Voss.

\hfill Carl-Johan Seger\parbox{1in}{\mbox{ }}

