\section{.sil format}

The .sil file format is reserved for SILOS II gate netlists.
In fact, the silos2exe converter supports only a quite small subset of the
SILOS II format.

To introduce the format, consider the following small example:
\begin{hol}
$********* GAte macros with delay parameters *******************
.macro CMOS4Xor2 i1 i2 / o
        o .xor 390,25.0 410,22.0 2.5*i1 2.5*i2
.eom

.macro CMOS4Nand2 i1 i2 / o
        o .nand 150,19.0 210,22.0 2.0*i1 2.0*i2
.eom

.macro CMOS4Nand3 i1 i2 i3 / o
        o .nand 180,21.0 240,24.0 2.0*i1 2.0*i2 2.0*i3
.eom


$********* Main circuit *******************
.TITLE ADDER   (On: Mon Feb  8 16:57:12 1993
(X1 CMOS4Xor2 a b t1
(X2 CMOS4Xor2 t1 cin result
(n1 CMOS4Nand2 a b t2
(n2 CMOS4Nand2 a cin t3
(n3 CMOS4Nand2 cin b t4
(n4 CMOS4Nand3 t2 t3 t4 cout

$********* End of netlist file *******************
\end{hol}

First, the .xor, and .nand commands refer to two of the 9 built-in
function types.
The numbers after the .xor component represent the rise-delay
and the fall-delay of the component and is calculated
as the first number plus the second number times the fanout load.
The multiplicative factors in front of input signals are used to
denote the load factor.
The macro definitions can be nested.
However, the scoping rules are not well defined (silos2exe uses dynamic
scoping), and thus it is recommended that all macros have distinct names.
Finally, the format of a macro call is: (instance-name name-of-macro
followed by arguments.

\subsection{Syntax of .sil format supported by silos2exe}

\begin{hol}
line            : alias\_line 
                | macro\_def 
                | component 
                | macro\_call 
                | 

alias\_line      : $ string = string

macro\_def       : .macro string {string}+ 
                           {line}*
                  .eom

macro\_call      : ( string string {input}+

component       : string .buf delays {input}+
                | string .clk number string
                | string .or delays {input}+
                | string .nor delays {input}+
                | string .nand delays {input}+
                | string .and delays {input}+
                | string .xor delays {input}+
                | string .inv delays {input}+
                | string .tbuf delays {input}+

delays          : delay delay
                | /* empty */

delay           : number
                | number , number
                | number , float

input           : load string
                | load - string

load            : number *
                | float *
                | /* empty */

\end{hol}


