\begin{thebibliography}{99}

\bibitem{beatty-mit90}
D. L. Beatty, R. E. Bryant, and C.-J. H. Seger,
``Synchronous Circuit Verification by Symbolic Simulation: An Illustration,''
{\it Sixth MIT Conference on Advanced Research in VLSI},
1990, pp.~98--112.

\bibitem{beatty-tr93}
D. L. Beatty, ``A Methodology for Formal Hardware Verification, with
Application to Microprocessors,'' Technical Report CMU-CS-;93-190,
Carnegie Mellon University, 1993.

\bibitem{beatty-dac94}
D. L. Beatty, and R. E. Bryant,
``Formally Verifying a Microprocessor using a Simulation Methodology,''
{\em 31st Design Automation Conference},
June, 1994, pp.~596--602.


\bibitem{bose-iccd89}
S. Bose, and A. L. Fisher,
``Verifying Pipelined Hardware Using Symbolic Logic Simulation,''
{\it International Conference on Computer Design},
IEEE, 1989, pp.~217--221.

\bibitem{bose-imec89}
S. Bose, and A. L. Fisher,
``Automatic Verification of Synchronous Circuits using Symbolic Logic
Simulation and Temporal Logic,''
{\it IMEC-IFIP International Workshop on Applied Formal Methods for Correct
VLSI Design}, 1989, pp.~759--764.

\bibitem{bryant-ieeetc86}
R. E. Bryant,
``Graph-Based Algorithms for Boolean Function Manipulation'',
 {\it IEEE Transactions on Computers}, Vol.\ C-35, No.\ 8 (August, 1986),
pp.~677--691.

\bibitem{bryant-dac87}
R. E. Bryant, D. Beatty, K. Brace, K. Cho, and T. Sheffler,
``COSMOS: a Compiled Simulator for MOS Circuits,''
{\it 24th Design Automation Conference}, 1987,
9--16.

\bibitem{bryant-tcad87b}
R. E. Bryant,
``Boolean Analysis of MOS Circuits,''
{\it IEEE Transactions on Computer-Aided Design of Integrated Circuits and
Systems},
Vol.~CAD-6, No.~4 (July, 1987), 634--649.

\bibitem{bryant-cav90}
R. E. Bryant, and C.-J. H. Seger, ``Formal Verification of Digital
Circuits Using Symbolic Ternary System Models,'' {\it Computer-Aided
Verification '90}, E. M. Clarke, and R. P. Kurshan, {\em eds.}
American Mathematical Society, 1991, pp.~121--146.

\bibitem{bryant-tcom91}
R. E. Bryant,
``On the Complexity of VLSI Implementations and Graph Representations of
Boolean Functions with Application to Integer Multiplication,''
{\it IEEE Transactions on Computers}, Vol.~40, No.~2 (February, 1991),
pp.~205--213.

\bibitem{bryant-tcad91}
R. E. Bryant,
``Formal Verification of Memory Circuits by Switch-Level Simulation,''
{\it IEEE Transactions on Computer-Aided Design
of Integrated Circuits and Systems}, Vol.~10, No.~1 (January, 1991),
pp.~94--102.

\bibitem{bryant-jacm91}
R. E. Bryant, ``A Methodology for Hardware Verification Based on Logic
Simulation,'' {\it J.ACM}, Vol.~38, No.~2 (April, 1991), pp.~299--328.

\bibitem{bryant-dac91}
R. E. Bryant, D. E. Beatty, and C.-J. H. Seger,
``Formal Hardware Verification by Symbolic Ternary Trajectory
Evaluation,''
{\it 28th Design Automation Conference}, June, 1991, pp.~297--402.


\bibitem{brzozowski-ieeetc79}
J. A. Brzozowski, and M. Yoeli.
``On a Ternary Model of Gate Networks.''
{\it IEEE Transactions on Computers C-28}, 3 (March 1979),
pp.~178--183.

\bibitem{burch-dac90}
J. R. Burch, E. M. Clarke, K. L. McMillan, and D. L. Dill,
``Sequential Circuit Verification Using Symbolic Model Checking,''
{\it 27th Design Automation Conference}, 1990, pp.~46--51.

\bibitem{clarke-toplas86}
E. M. Clarke, E. A. Emerson, and A. P. Sistla,
``Automatic Verification of Finite-State Concurrent Systems Using Temporal
Logic Specifications,''
{\em ACM Transactions on Programming Languages},
Vol.~8, No.~2 (April, 1986), pp.~244--263.

\bibitem{clarke-popl92}
E. M. Clarke, O. Grumberg, and D. E. Long,
``Model Checking and Abstraction,''
{\em Proc.~19th Annual ACM Symposium on Principles of Programming Languages},
Jan., 1992.

\bibitem{coudert-imec89}
O. Coudert, C. Berthet, and J. C. Madre,
``Verification of Sequential Machines using Boolean Functional Vectors,''
{\it IMEC-IFIP International Workshop on Applied Formal Methods for Correct
VLSI Design}, 1989, pp.~111--128.

\bibitem{coudert-cav90}
O. Coudert, J.-C. Madre, and C. Berthet,
``Verifying temporal properties of sequential machines without
building their state diagrams,''
{\it Computer-Aided Verification `90}, E. M. Clarke, and
R. P. Kurshan, {\em eds}.
American Mathematical Society, pp.~75--84.

\bibitem{darringer-dac79}
J. A. Darringer,
``The Application of Program Verification Techniques to Hardware
Verification,''
{\it 16th Design Automation Conference}, 1979, pp.~375--381.

\bibitem{devadas-dac87}
S. Devadas, H.-K. T. Ma, and A. R. Newton,
``On the Verification of Sequential Machines at Differing Levels of
Abstraction,''
{\it 24th Design Automation Conference}, 1987, pp.~271--276.

\bibitem{kam-iccd92}
T. Kam, and P. A. Subramanyam,
``Comparing Layouts with HDL Models: A Formal Verification Technique,''
{\it International Conference on Computer Design},
IEEE, 1992, pp.~588--591.


\bibitem{LCF}
M. Gordon, R. Milner, and C. Wadsworth, ``Edinburgh LCF'',
{\em Lecture Notes in Computer Science},
No.~78, Springer Verlag, 1979.

\bibitem{gordon-favd85}
M. Gordon,
``Why higher-order logic is a good formalism for specifying and verifying
hardware,''
{\em Formal Aspects of VLSI Design}, G. Milne and P. A. Subrahmanyam, {\em
eds.}, North-Holland, 1986, pp.~153--177.

\bibitem{HOLsystem}
Michael J. C. Gordon et~al., {\em The HOL System Description},
Cambridge Research Centre, SRI International, Suite~23, Miller's Yard,
Cambridge CB2 1RQ, England.

\bibitem{HOLbook}
M.J.C. Gordon and T.F. Melham (eds.),
``Introduction to HOL'',
Cambridge University Press, 1993.


\bibitem{halbwachs-tse92}
N. Halbwachs, F. Lagnier, and C. Ratel,
``Programming and Verifying Real-Time Systems by Means of the Synchronous Data-F
low Language LUSTRE,''
{\em IEEE Transactions on Software Engineering},
Vol.~18, No.~9 (September, 1992), pp.~785--793.

\bibitem{JoyceGeneric}
J. Joyce, ``Generic Structures in the Formal Specification and Verification
of Digital Circuits'', {\em The Fusion of Hardware Design and Verification},
G. Milne, {\em ed.}, North Holland, 1988, pp.~50--74.

\bibitem{JoyceSegerDAC}
J. Joyce and C. Seger, ``Linking BDD-Based Symbolic Evaluation
to Interactive Theorem-Proving'', {\it 30th Design Automation Conference},
1993, pp.~469--474.

\bibitem{jephson-69}
J. S. Jephson,  R. P. McQuarrie, and R. E. Vogelsberg,
``A Three-Level Design Verification System,''
{\it IBM Systems Journal} Vol.~8, No.~3
(1969), pp.~178--188.

\bibitem{kurshan-tcad91}
R. P. Kurshan, and K. L. McMillan,
``Analysis of Digital Circuits Through Symbolic Reduction,''
{\em IEEE Transactions on Computer-Aided Design
of Integrated Circuits and Systems}, Vol.~10, No.~11 (November, 1991),
pp.~1356--1371.

\bibitem{mead-80}
C. A. Mead, and L. Conway,
{\it Introduction to VLSI Systems},
Addison-Wesley, 1980.

\bibitem{ML}
R. Milner, ``A Proposal for Standard ML'',
{\it Proceedings of ACM Conference on LISP and Functional Programming},
Austin, TX, Aug. 1984, pp.~184--197

\bibitem{pnueli-focs77}
A. Pnueli,
``The Temporal Logic of Programs,''
{\em 18th Symposium on the Foundations of Computer Science},
IEEE, 1977, pp.~46--56.

\bibitem{reeves-tcad87}
D. S. Reeves, and M. J. Irwin,
``Fast Methods for Switch-Level Verification of MOS Circuits'',
{\it IEEE Transactions on CAD/IC},
Vol.~CAD-6, No.~5 (Sept., 1987), pp.~766--779.

\bibitem{seger-imec89}
C-J. Seger, and R. E. Bryant,
``Modeling of Circuit Delays in Symbolic Simulation'',
{\it IMEC-IFIP International Workshop on Applied Formal Methods for Correct
VLSI Design}, 1989, pp.~625--639.

\bibitem{SegBry95}
C-J. Seger, and R. E. Bryant, ``Formal Verification by Symbolic Evaluation
of Partially-Ordered Trajectories'', {\it Formal Methods in System Design},
Vol. 6, March 1995, pp. 147--189, 1995.

\bibitem{SegerJoyceTwoLevel}
C. Seger and J. Joyce, ``A Mathematically Precise Two-Level Formal Hardware
Verification Methodology'', Technical Report 92-34, Department of Computer
Science, University of British Columbia, December 1992.


\bibitem{Stoy}
J. Stoy, {\it Denotational Semantics: The Scott-Strachey Approach to
Programming Language Theory}, MIT Press, 1977.

\bibitem{Miranda}
D. Turner, ``MIRANDA: A Non Strict Functional Language with Polymorphic Types'',
in {\it Proceedings of the IFIP International Conference on Functional Programming Languages and Computer Architecture}, Lecture Notes in Computer Science,
Springer Verlag, Vol.~201, 1985.

\bibitem{wolper-popl86}
P. Wolper,
``Expressing Interesting Properties of Programs in Propositional Temporal Logic,
''
{\em Proc.~13th Annual ACM Symposium on Principles of Programming Languages},
Jan., 1986, pp.~184--193.

\end{thebibliography}

