\section{Fl---The Meta Language of Voss}

In this section\footnote{This chapter is to a
large extent modeled after Chapter 1 in the HOL System DESCRIPTION
from Cambridge University. In particular, many of the early examples are
taken from this source.} we provide an introduction to
the functional language FL.

Similar to many theorem provers (e.g., the HOL system[HOL]) the
Voss command language for the verification system is a general
purpose language. In fact, it shows a strong degree of similarity
to the version of ML used in the HOL system. However, there are
several differences:  many syntactic but some more fundamental. In
particular, the functional language used in Voss has lazy evaluation
semantics. In other words, no expression is evaluated until it is
absolutely needed. Similarly, no expression is evaluated more than
once. Another difference is that Boolean functions are first-class
objects and can be created, evaluated, compared and printed out.
For efficiency reasons these Boolean functions are represented as
ordered binary decision diagrams. 

Fl is an interactive language. At top-level one can:
1) perform declarations, and 2) evaluate expressions.
In this section we will introduce the language by several examples.

\subsection{Invoking fl}

If the Voss system is installed on your system
and you have the suitable search path set up, it suffices
to type fl to get a stand-alone version of Fl. In this manual, we
have used Voss 1.5 throughout. In other words, typing fl yielded:

\begin{hol}
% fl
      /\verb@\@
     /  \verb@\@/\verb@\@
 /\verb@\@ /      \verb@\@
/  Voss 1.5 \verb@\@
VOSS-LIBRARY-DIRECTORY = /isd/local/generic/lib/vosslib

:
\end{hol}
Note that the VOSS-LIBRARY-DIRECTORY is installation dependent.
We will return to this later when we discuss the user defaults.

The fl program can take a number of arguments.
In particular, 
\begin{description}
\item[-f n]
Start FL by first reading in the content of the file named n.
\item[-I n]
Set the default search path to n.
\item[-s i]
Set the default OBDD table to be of size $2^i$, where $i$ can range
from 16 to 22.
Normally, this is not needed.
However, if a verification task will be needing more than 1/2 million
OBDD nodes, setting $i$ to some number above 19 will improve performance.
\end{description}

\subsection{Expressions}
The Fl prompt is : so lines beginning with this contain
the user's input; all other lines are output of the system.
\begin{hol}
: 2+3;
5
\end{hol}
Here we simply evaluated the expression 2+3 and FL reduced
it to normal form; in this case computed the result 5. Note that
fl does only support integers as number types. Furthermore, these
integers are limited to +/- 536870912 (two's complement 30 bit
numbers). This restriction on the numbers is likely to disappear
shortly, but for Voss 1.5 it is a restriction that is important to
remember.

\subsection{Declarations}

The declaration let x = e binds a computation of e to
the variable x. Note that it does not evaluate e (since the language
is fully lazy). Only if x is printed or used in some other expression
that is evaluated will it be evaluated. Also, once e is evaluated,
x will refer to the result of the evaluation rather than the
computation. Hence, the expression e is evaluated at most once,
but it may not be evaluated at all.
\begin{hol}
: let x = 3+3;
x::int
\end{hol}
Note
that when expressions are bound to variables, the system simply
prints out the inferred type of the expression. We will return to
the typing scheme in FL later. For now, it suffices to say that FL
tries to find as general type as possible that is consistent with
the type of the expression.

Contrary to ML, FL 1.5 does not allow
simultaneous bindings. Hence, if we would like to bind the expressions
2 and 4-5 to the variables x and y respectively, we would have to
write:
\begin{hol}
: let x = 2;
x::int
: let y = 4-5;
y::int
\end{hol}
A declaration
can be made local to the evaluation of an expression e by evaluating
the expression decl in e. For example:
\begin{hol}
: let y = let x = 4 in x-5;
y::int
\end{hol}
would bind the expression 4 to x only inside the expression
bound to y. Thus, we get:
\begin{hol}
: let x = 2;
x::int
: let y = let x = 4 in x-5;
y::int
: x;
2
: y;
-1
\end{hol}
FL is lexically scoped, and thus
the binding in effect at the time of definition is the one used.
In other words, if we write:
\begin{hol}
: let x = 2;
x::int
: let y = x*5;
y::int
: let x = 12;
x::int
\end{hol}
and we then evaluate y we will get 10
rather than 60.

\subsection{Functions}

To define a function f with formal parameter x and body e one performs
the declaration:
let f x = e. To apply the function f to an actual parameter e one evaluates
the expression f e.
\begin{hol}
: let f x = x+2;
f::(int) -> (int)
: f 4;
6
\end{hol}
Note that the type inferred for f is essentially ``a function taking
an int as argument and
returning an int''. Applications binds more tightly than anything else
in FL; thus for example:
f 3 + 4 would be evaluated as: ((f 3)+4) and thus yield 9.

Functions of several arguments can also be defined:
\begin{hol}
: let add x y = x+2*y;
add::(int) -> ((int) -> (int))
: add 1 4;
9
: let f = add 1;
f::(int) -> (int)
: f 4;
9
\end{hol}
Applications associate to the left so add 3 4 means (add 3) 4. In the
expression add 3, the
function add is partially applied to 3; the resulting value is the
function of type int->int which
adds 3 to twice its argument. Thus add takes its arguments one at a
time. We could have made
add take a single argument of the cartesian product type (int\verb@#@int):
\begin{hol}
: let add (x,y) = x+y;
add::((int # int)) -> (int)
: add (3,4);
7
: add 3;
===Type mismatch: (int # int) and int
#### Run-time error
---- Type error
\end{hol}
As well as taking structured arguments (e.g. (3,4)) functions may also return
structured results:
\begin{hol}
: let manhat_dist (x1,y1) (x2,y2) = (x2-x1, y2-y1);
manhat_dist::((int # int)) -> (((int # int)) -> ((int # int)))
: manhat_dist (1,1) (3,5);
(2,4)
\end{hol}

Trying to print a function with insufficient number of actual arguments
yield a dash
for the function and the type of the expression is printed out. For example:
\begin{hol}
: (5, manhat_dist (1,2));
(5,-) ::(int # ((int # int)) -> ((int # int)))
\end{hol}

\subsection{Recursion}

The following is an attempt to define the factorial function:
\begin{hol}
: let fact n = n=0 => 1 | n*fact (n-1);
#### Run-time error
---- Undefined variable (fact)
\end{hol}
The problem is that any free variables in the body of a function have
the bindings they had
just before the function was declared; fact is such a free variable
in the body of the declaration
above, and since it is not defined before its own declaration, an error
results. To make things
clear consider:
\begin{hol}
: let f n = n+1;
f::(int) -> (int)
: let f n = n=0 => 1 | n*f (n-1);
f::(int) -> (int)
: f 3;
9
\end{hol}
Here 3 results in the evaluation of 3*(f 2), but now the first f is
used so f 2 evaluates to
2+1=3. To make a function declaration hold within its own body, letrec
instead of let must be
used. The correct recursive definition of the factorial function is thus:
\begin{hol}
: letrec fact n = n=0 => 1 | n*fact (n-1);
fact::(int) -> (int)
: fact 5;
120
\end{hol}

It should be pointed out that FL currently does not allow direct definition
of mutually recursive functions.
One way around this is to define a ``wrapper'' function that takes as a
parameter some number of name of the mutually recursive function that
is intended.
Mutually recursive function definitions is quite likely to be added to
the next major release of the system.

\subsection{Tuples}

If $e_1, e_2, \ldots , e_n$ have types $t_1, t_2, \ldots , t_n$, then
the FL expression $(e_1, e_2, \ldots , e_n)$ have type
$t_1$\verb@#@$t_2$\verb@#@$\ldots$\verb@#@$t_n$.
The standard functions on tuples are fst (first), snd (second), and the infix
operation , (pair).
\begin{hol}
: let q = ((1,2),3);
q::((int # int) # int)
: let qq = (1,2,3);
qq::(int # (int # int))
: q;
((1,2),3)
: qq;
(1,2,3)
: let qqq = (1,"abc");
qqq::(int # string)
: qqq;
(1,"abc")
\end{hol}

\subsection{Lists}

If $e_1, e_2, \ldots , e_n$ have type $t$, then the FL expression
$[e_1, e_2, \ldots , e_n]$ has type ($t$ list).
The standard functions on lists are hd (head), tl (tail), [] (the empty
list), and the infix operation :
(cons). Note that all elements of a list must have the same type (compare
this with a tuple where
the size is determined but each member of the tuple can have different type).
\begin{hol}
: let l = [1,2,3,3,2,1,2];
l::(int list)
: hd l;
1
: tl l;
[2,3,3,2,1,2]
: 0:l;
[0,1,2,3,3,2,1,2]
: letrec (len [] = 0) /\verb!\! (len (a:rest) = 1+len rest);
len::((* list)) -> (int)
: len l;
7
\end{hol}

\subsection{Strings}

A sequence of characters enclosed between " or ` is a string. The standard
functions on strings
are \^ (catenation), explode (make string into list of strings) and
implode (make list of strings into single string).
There are also int2str and bool2str functions that create a string
from an integer or an object of type boolean.
We will return to these later.
\begin{hol}
: let q = "abc and _12!@@#";
q::string
: let qq = `qw"q qw`;
qq::string
: q^qq;
"abc and _12!@@#qw"q qw"
: explode q;
["a","b","c"," ","a","n","d"," ","_","1","2","!","@","@","#"]
: implode ["1", "2"];
"12"
: int2str (1-34);
"-33"
: bool2str ((variable "a") AND (variable "b"));
"a&b"
\end{hol}

\subsection{Polymorphism}

The list processing functions hd, tl, etc. can be used on all types of lists.
\begin{hol}
: hd [1,2,3];
1
: hd ["abc", "edf"];
"abc"
: (hd ["a", "b"]), hd [4,2,1];
("a",4)
: let q = [T,T,F];
q::(bool list)
: hd q;
T
\end{hol}
Thus hd has several types; for example, it is used above with types
(int list) \verb!->! int,
(string list) \verb!->! string, and (bool list) \verb!->! bool.
In fact if ty is any type then hd has the type (ty list) \verb!->! ty.
Functions, like hd, with many types are called polymorphic,
and FL uses type variables *, **, ***, etc. to represent their types.
\begin{hol}
: let f x = hd x;
f::((* list)) -> (*)
: letrec map fn [] = []
   /\verb!\!    map fn (h:rest) = (fn h) : (map fn rest);
map::((*) -> (**)) -> (((* list)) -> ((** list)))
: letrec fact n = n=0 => 1 | n*fact (n-1);
fact::(int) -> (int)
: map fact [1,2,3,4,5,6,7];
[1,2,6,24,120,720,5040]
\end{hol}
The FL function map takes a function f (with argument type * and result
type **), and a list l
(of elements of type *), and returns the list obtained by applying
f to each element of l
(which is a list of elements of type **). Map can be used at any instance
of its type: above, both *
and ** were instantiated to int; below, * is instantiated to (int list)
and ** to bool. Notice that the
instance need  not be  specified; it is determined by the type checker.
\begin{hol}
: let eq1 x = x=1;
eq1::(int) -> (bool)
: map eq1 [1,2,3,4,12,2,1,2];
[T,F,F,F,F,F,T,F]
\end{hol}

It should be pointed out that FL has a polymorphic type system that
is sightly different from
standard ML's. In particular, only ``top-level'' user-defined functions
can be polymorphic. In
other words, the following works as we would expect.
\begin{hol}
: let null l = l = [];
null::((* list)) -> (bool)
: let f x y = null x OR null y;
f::((* list)) -> (((** list)) -> (bool))
: f [1,2,3] ["abc", "cdef"];
F
\end{hol}
However, if we use the same declaration inside the expression, it must
be monomorphic. In other
words, the following example fails.
\begin{hol}
: let f x y =
    let null l = l = [] in
    null x OR null y;
f::((* list)) -> (((* list)) -> (bool))
: f [1,2,3] ["abc", "cdef"];
===Type mismatch: int and string
#### Run-time error
---- Type error
\end{hol}
In this respect, FL is similar to the functional language called
Miranda\footnote{Miranda is a trademark.}\cite{Miranda}.

\subsection{Lambda Expressions}

The expression \verb@\@x.e evaluates to a function with formal parameter x
and body e. Thus the
declaration let f x = e is equivalent to let f = \verb@\@x.e. The character
\verb@\@ is our representation of lambda,
and expressions like \verb@\@x.e are called lambda-expressions.
\begin{hol}
: \verb!\!x.x+1;
- ::(int) -> (int)

: let q = \verb!\!x.x+1;
q::(int) -> (int)
: q 1;
2
: map (\verb!\!x.x*x) [1,2,3,4,5];
[1,4,9,16,25]
\end{hol}

\subsection{Failures}

Some standard functions fail at run-time on certain arguments, printing
out a string (which is usually the function name) to identify the sort
of failure. A failure with string "t" may also be
generated explicitly by evaluating the expression error "t" (or more
generally error e where e has type string).
\begin{hol}
: hd(tl [2]);
Failure:    ----  Cannot compute hd of the empty list

: 1/0;
Failure:    ----  Division by zero


: error "My message";
Failure:    ----   My message
\end{hol}

A failure can be trapped by catch; the value of the expression $e_1$ catch
$e_2$ is that of $e_1$, unless
$e_1$ causes a failure, in which case it is the value of $e_2$. One important
property of catch is that it is
(very) strict in its first argument. In other words, (hd ($e_1$ catch
$e_2$)) will completely evaluate $e_1$
even though only the first element in the list may be needed. In view
of FL's lazy semantics, the
use of catch should be very carefully considered. In particular, the
bindings of catch is dynamic
rather than static so the user beware! It is not unlikely that catch
will disappear from FL in future
versions.

\subsection{Boolean Expressions}

All Boolean expressions in FL are maintained as ordered binary decision
diagrams. Hence, it
is very easy to compare complex Boolean expressions and to combine
them in different ways.
Boolean variables are created by variable s, where s is of type string.
The system uses name
equivalence, and thus
\begin{hol}
: let v = variable "v";
v::bool
: v=v;
T
: variable "v" = variable"v";
T
\end{hol}

The constants true and false are denoted T and F respectively. The
standard boolean functions
are available, i.e., AND, OR, NOT, XOR, and = are all defined for objects of type Boolean..
Furthermore, there
is a special identity operator == that return true or false depending
on whether the two arguments
represent the same Boolean function or not.

Note that the variable ordering in the OBDD representation is defined
by the order in which each variable function call {\em gets evaluated}.
Since FL is a fully lazy language, and thus the order in which expressions
are evaluated is often difficult to predict, it is strongly recommended
that each variable declaration is forced to be evaluated before it
is being used.
In the standard library default.fl a function, called declare, is
defined to simplify this task.
We will return to this later when we discuss the various FL libraries.
Also, note that once a variable function call has been evaluated for
a specific string argument, the created variable has been placed
in the variable order and thus consequent calls will return this
variable.
Consequently, the only way of changing the variable order after a variable
has been created, is to quit FL and start it again.

\begin{hol}
: let a = variable "a";
a::bool
: let b = variable "b";
b::bool
: a AND b;
a&b
: a OR b;
a + b
: NOT a AND NOT b AND T;
a'&b'
: a = b;
a&b + a'&b'
:  a == b;
F
: (a=b) == (a AND b OR NOT a AND NOT b);
T
\end{hol}

The default style for printing Boolean expressions is as a sum-of-products.
Since this may
require printing an extremely large expression, there is a user-setable
limit on how many products
that will be printed and the maximum size of a product. For more details
how to modify these two
parameters, see the section on the .vossrc file on page~\ref{vossrc}.

\subsection{Quantifiers}

There are several ways of using quantification. But the ``traditional''
!x. e (for all x) and ?x. e
(there is an x) can be used as long as the type of x and e is bool.
In addition, you can also
quantify away a variable in an expression by quant\_forall v e or quant\_thereis
v e.
\begin{hol}
: !a. ?b. (a XOR b);
T
: let a = variable "a"; let b = variable "b"; let c = variable "c";
a::bool
b::bool
c::bool
: : : a AND b OR c;
a&b + c
:  quant_forall a (a AND b OR c);
c
: quant_thereis (a OR c) (a AND b OR c);
T
\end{hol}
In fact, quant\_forall and quant\_thereis quantifies away all variables
in the first Boolean
expression. For example:
\begin{hol}
: let v s = variable s;
v::(string) -> (bool)
: let a = v "a"; let b = v "b"; let c = v "c"; let d = v "d";
a::bool
b::bool
c::bool
d::bool
: : : : let ex = (a AND NOT b);
ex::bool
: ex;
a&b'
: let nex = ex AND (a=c) AND (b=d);
nex::bool
: quant_thereis (a AND b) nex;
c&d'
\end{hol}
Note that the actual Boolean expression used as first argument is irrelevant.
The only important fact is on what variables the expression depends.

\subsection{Dependency}

Sometimes it is useful to find out which Boolean variables a
Boolean function actually depends on.
The built-in function depends takes a list of elements of type
bool and return the union of the variables these functions depend on.
For example:
\begin{hol}
: let v s = variable s;
v::(string) -> (bool)
: let a = v "a"; let b = v "b"; let c = v "c"; let d = v "d";
a::bool
b::bool
c::bool
d::bool
: : : : let ex1 = (a=c) AND d;
ex1::bool
: let ex2 = a = b;
ex2::bool
: depends [ex1];
["a","c","d"]
: depends [ex1,ex2];
["a","b","c","d"]
\end{hol}

Note that the order of the variables in the list returned by depends
is the variable order of the OBDD representation.

\subsection{Substitutions}

Given a Boolean function represented as a OBDD, it is convenient to
be able to apply the
function to some arguments. This can be accomplished by the substitute
command that takes a
list of (variable name, expression) and an expression in which the
simultaneous substitution is to
be made. For example,:
\begin{hol}
: let v s = variable s;
v::(string) -> (bool)
: let a = v "a"; let b = v "b"; let c = v "c"; let d = v "d";
a::bool
b::bool
c::bool
d::bool
: : : : let ex = (a AND NOT b);
ex::bool
: ex;
a&b'
: substitute [("a", c), ("b", d)] ex;
c&d'
\end{hol}

It should be pointed out that there are no restrictions on the expressions
in the substitutions.
In particular, it is possible to ``swap'' variables. We illustrate this by continuing the example above:
\begin{hol}
: ex;
a&b'
: substitute [("a", b), ("b", a)] ex;
a'&b
\end{hol}

\subsection{Type Abbreviations}

Types can be given names:
\begin{hol}
: new_type_abbrev pair = int#int;
: let p = (1,2);
p::(int # int)
\end{hol}
However, as can be seen from the example, the system does not
make any distinction between the new type name
and the actual type. It is purely a short hand that is useful when
defining concrete types below.

\subsection{Concrete Types}

New types (rather than mere abbreviations) can also be defined. Concrete
types are types
defined by a set of constructors which can be used to create objects
of that type and also (in
patterns) to decompose objects of that type. For example, to define
a type card one could use the
construct type:
\begin{hol}
: lettype card = king | queen | jack | other int;
other::(int) -> (card)
jack::card
queen::card
king::card
\end{hol}
Such a declaration declares king, queen, jack and other as constructors
and gives them
values. The value of a 0-ary constructor such as king is the constant
value king. The value of a
constructor such as other is a function that given an integer value
n produces other(n).
\begin{hol}
: king;
- ::card

: other 9;
- ::card
\end{hol}

Note that there is no print routine for concrete types. If a print
routine is desired, one has to
define it.
To define functions that take their argument from a concrete type,
we introduce the idea of pattern matching.
In particular
\begin{hol}
let f pat1 = e1
 /\verb@\@ f pat2 = e2
 /\verb@\@ ...
 /\verb@\@ f patn = en;
\end{hol}
denotes a function that given a value v selects the first pattern that
matches $v$, say pati, binds the variables of pati to
the corresponding components of the value $v$ and then evaluates the
expression ei.
We could for example define a print function for the cards in the following way:
\begin{hol}
: let pr_card king = "K"
  /\verb@\@  pr_card queen = "Q"
  /\verb@\@  pr_card jack = "J"
  /\verb@\@  pr_card (other n) = int2str n;
pr_card::(card) -> (string)
: pr_card king;
"K"
: pr_card queen;
"Q"
: pr_card jack;
"J"
: pr_card (other 5);
"5"
\end{hol}

Mutually recursive types can also be defined. To do so, simply list
the type names on the left
hand side of the equality sign and list the type expressions on the
right hand side. For example:
\begin{hol}
: lettype IExpr, BExpr = Ivar string | Plus IExpr IExpr,
                         And BExpr BExpr | GEQ IExpr IExpr;
GEQ::(IExpr) -> ((IExpr) -> (BExpr))
And::(BExpr) -> ((BExpr) -> (BExpr))
Plus::(IExpr) -> ((IExpr) -> (IExpr))
Ivar::(string) -> (IExpr)
\end{hol}

\subsection{Abstract Types}

In FL one can also hide the definitions of types, type constructors,
and functions. By
enclosing a sequence of type declarations and function definitions
within begin\_abstype
end\_abstype elist, only the constructors and/or functions mentioned
in the elist will be visible
and accessible for other functions and definitions. Thus, one can protect
a concrete type and only
make some abstract constructor functions available. To illustrate the
concept, consider defining a
concrete type called theorem. The only way we would like the user to
be able to create a new theorem is to give a Boolean expression that
denotes a tautology (something always true).
First we define the expression type.
\begin{hol}
: lettype expr =  Forall string expr |
                Thereis string expr |
                Var string |
                True |
                False |
                And expr expr |
                Or expr expr |
                Not expr;
Not::(expr) -> (expr)
Or::(expr) -> ((expr) -> (expr))
And::(expr) -> ((expr) -> (expr))
False::expr
True::expr
Var::(string) -> (expr)
Thereis::(string) -> ((expr) -> (expr))
Forall::(string) -> ((expr) -> (expr))
\end{hol}

We then define the concrete type theorem and the constructor function
is\_taut. Note that we
also define a couple of help functions. However, only the is\_taut function
is exported out of the
abstract type, and thus is the only way of creating a theorem.
\begin{hol}
: begin_abstype;
: lettype theorem = Thm expr;
Thm::(expr) -> (theorem)
: letrec assoc x l = l = [] => error "assoc" |
                let h = hd l in
                (fst h) = x => (snd h) | assoc x (tl l);
assoc::(*) -> ((((* # **) list)) -> (**))
: letrec  (eval (Forall s e) al = !x.(eval e ((s,x):al))) /\verb@\@
        (eval (Thereis s e) al = ?x.(eval e ((s,x):al))) /\verb@\@
        (eval (Var s) al = (assoc s al) catch
                                   (error "Free variable")) /\verb@\@
        (eval True al = T) /\verb@\@
        (eval False al = F) /\verb@\@
        (eval (And e1 e2) al = (eval e1 al) AND (eval e2 al))/\verb@\@
        (eval (Or e1 e2) al = (eval e1 al) OR (eval e2 al)) /\verb@\@
        (eval (Not e) al = NOT (eval e al));
eval::(expr) -> ((((string # bool) list)) -> (bool))
: let is_taut e = ((eval e []) == T) => Thm e |
                                       error "Not a tautology";
is_taut::(expr) -> (theorem)
end_abstype is_taut;
\end{hol}

We can now use this very safe theorem system, since we can only generate
theorems that are
tautologies. For example
\begin{hol}
: let e = (Forall "a" (Thereis "b" (And (Var "a") (Var "b"))));
e::expr
: is_taut e;
Failure:    ----   Not a tautology


: let f = (Forall "a" (Thereis "b" (Or (Var "a") (Var "b"))));
f::expr
: is_taut f;
- ::theorem
\end{hol}

\subsection{Infix Operators}

In order to make the FL code more readable, it is possible to declare
a function to be infix
(associating from the left), infixr (associating from the right), nonfix
(really prefix), or postfix.
For the infix and infixr directives, the precedence can be given as
a number from 1 to 9, where a
higher number binds tighter. Note that prefix and postfix functions
bind higher than any infix
function. Beware that the fixity declaration modifies the parser and
thus remains in effect whether
the function is exported out of an abstract data type or note. This
``feature'' is likely to be fixed
fairly soon. As an illustration of this idea, consider the following example:
\begin{hol}
: lettype expr = Var int |
                Mult expr expr |
                Plus expr expr |
                Negate expr;
Negate::(expr) -> (expr)
Plus::(expr) -> ((expr) -> (expr))
Mult::(expr) -> ((expr) -> (expr))
Var::(int) -> (expr)
: letrec eval (Var i) = i /\verb@\@
         eval (Mult e1 e2) = (eval e1) * (eval e2) /\verb@\@
         eval (Plus e1 e2) = (eval e1) + (eval e2) /\verb@\@
         eval (Negate e1) = 0-(eval e1);
eval::(expr) -> (int)
: let ** a b = Mult a b;
**::(expr) -> ((expr) -> (expr))
: let ++ a b = Plus a b;
++::(expr) -> ((expr) -> (expr))
: infix 4 **;
: infix 3 ++;
: let ' i = Var i;
'::(int) -> (expr)
: let q = '1 ++ Negate ('2) ** Negate ('4);
q::expr
: eval q;
9
\end{hol}

The next example illustrates how postfix declarations can make the
code more readable.
\begin{hol}
: let ns i = 1000*i;
ns::(int) -> (int)
: postfix ns;
: let to a b = (a,b);
to::(*) -> ((**) -> ((* # **)))
: infix 3 to;
: 2 ns to 4 ns;
(2000,4000)
\end{hol}

\subsection{Circuit Models}

Since the main use of the FL system, and its historical root, is related
to hardware verification, there are a number of built-in
functions specifically tailored towards hardware modeling
and symbolic trajectory evaluation in particular. 
Internally, a circuit is represented by a list of nodes (names),
and a next state function.
The next state function is mapping the current state of the circuit
(including the current values on the inputs) to a new state
of the circuit.
Since the circuit representation is intended for trajectory evaluation,
the value domain for each node in the circuit is $\{0,1,\X,\top\}$
and thus the next state function consists of quaternary (four-valued)
extensions of the usual Boolean function.
The type of such circuit is fsm (for finite state machine) and, by
default, it does not have a print function (since the machines
are usually much too large to be meaningful anyway to the user.
However, for the curious reader, the command print\_fsm will print out
a pretty complete version of the state machine.
Note that for efficiency reasons, the next state function also
contains delay and fan-in and fan-out information to aid in the
efficient simulation.

In general, there are two ways of creating a fsm object:
\begin{enumerate}
\item
Loading a pre-compiled version of a circuit in .exe format.
\item
Converting a FL structure into an fsm.
\end{enumerate}

\subsubsection{Loading a .exe File}

If the original circuit was described in Berkeley .sim format or
as a SILOS II gate list, there are programs distributed with the
system that can be used to compile an fsm model directly from
these formats.
The common format for such pre-compiled circuit model is a (binary)
file with a .exe suffix.
%In Section~\ref{CompanionTools} we discuss in more detail how
%to create such .exe file.
Loading in the .exe file and making it an fsm object simply
involves calling the load\_exe function.

For example, if there is a full\_adder.exe file in the current
working directory, the following command would create an fsm object
of the circuit.
\begin{hol}
: let ckt1 = load_exe "full_adder.exe";
ckt1::fsm
\end{hol}
Note that there is no theoretical limit on the number of circuit that can
be loaded into the FL system at any particular point in time.
However, since fsm models are often quite large, it is generally
advisable not to load more models than absolutely necessary.
%(In most cases, only one model will be used, but there are times
%when several models can be useful. We will return to this later
%in Section~\ref{ModelChecking}.)

\subsubsection{Creating an fsm Object Inside FL}

{\bf Warning:} This addition is fairly young and has thus not
been extensively tested. Also, it leaves quite a bit to
be desired in terms of not being very ``clean and simple''.
For a reader that is more interested in using the Voss system rather
than writing a new conversion program from some other netlist format, this
section can be skipped.

The main routine for creating an fsm object is make\_fsm.
The type of make\_fsm is (Set) -> (fsm)\footnote{Currently, when
FL is invoked, the type of make\_fsm is actually (*)\verb!->!(fsm), but
this is a bug that will be removed shortly.}
Before going into how an object of type Set can be constructed, we
need to make a small digression and introduce the quaternary logic type.
For historical reasons, the name of the quaternary objects is tern,
and tern is defined as the concrete type:
\begin{hol}
: lettype tern =  One |
                Zero |
                X |
                Z |
                Val string |
                And tern tern |
                Or tern tern |
                Not tern;
Not::(tern) -> (tern)
Or::(tern) -> ((tern) -> (tern))
And::(tern) -> ((tern) -> (tern))
Val::(string) -> (tern)
Z::tern
X::tern
Zero::tern
One::tern
\end{hol}
where One, Zero, X, And, Or, and Not are the obvious functions.
Z is used to represent the top value.
Finally, Val s is used to refer to the value on node s.
Thus, as a concrete example, the next state function of a node "o" may
be described as (Not (And (Val "in1") (Val "in2"))).

There are four constructor functions for an object of type Set:
\begin{enumerate}
\item
Empty,
\item
Element (string\verb@#@((tern\verb@#@tern) list)),
\item
Union Set Set, and
\item
Sequential Set Set.
\end{enumerate}
The Empty is introduced to make writing functions that generate
objects of type Set easier.
Element is the constructor that actually introduces new nodes and
also give driver functions for the node.
In general, Element n gvl, will create a node named n.
This node will at least (more about this later) the drivers
listed in the gvl list.
A driver is simply a pair of quaternary expressions:
the first object of the pair being a guard, and the second part
the value being driven when the guard is true.
For a combinational node, the driver list is a single pair whose
guard is One and whose value component is the next state function
of the gate.
For more complex nodes, like register nodes or bus nodes, the guard-value
list is often much longer.

The Union construct is used to gather a collection of these Element
declarations so that an fsm can eventually be made.
Finally, the Sequential constructor takes two objects S1 and S2 of
type Set and merges them into a single object of type Set.
In that sense, Sequential behaves pretty much like Union.
The difference is how the two deal with nodes defined in both set S1 and set S2
and for which both nodes may have at least one of their guards enabled
at the same time.
Union will find the greatest lower bound of the values being driven
at the same time whereas Sequential will assume that the driver in the
second set overrides the driver in the first.
To illustrate the difference, consider the following example:
\begin{hol}
: let a = Element ("in1", [(One, One)]);
a::Set
: let b = Element ("in1", [(One, Zero)]);
b::Set
: let ex1 = Union a b;
ex1::Set
: let ex2 = Sequential a b;
ex2::Set
: let m1 = make_fsm ex1;
m1::fsm
: let m2 = make_fsm ex2;
m2::fsm
: m1;
\end{hol}
Now m1 and m2 will both be fsm models with a single node called in1.
However, in m1, the next state function of in1 will be $\X$, since that
is the most we can say from the inconsistent driver commands given
in the example.
On the other hand, the next state function of the in1 node in m2 will be Zero.

For an example of using this facility, the directory HDL in the demo
distribution illustrates how one can go about defining a new description
language in order to create an fsm object.

\subsection{Queries to the Circuit Model}

There are four built-in functions that are useful in dealing with
an fsm model: nodes, fanin, and fanout, and get\_node\_val.
Nodes takes an fsm model and returns a {\em list of lists} of circuit nodes.
The reason for the double listing is that a node may have more than
one name (only from .exe file translations).
Fanin takes an fsm model and a name of a node and returns the
list of node names that the next state function depend on.
Fanout works in a similar way, but for the fanout nodes.
Finally, get\_node\_val takes an fsm object and the name of a node
and returns the encoded version of the current value of the node.
The encoding used is: $\X = (T,T)$, $1=(T,F)$, $0=(F,T)$, and $\top=(F,F)$.
Of course, the two Boolean values are often Boolean functions over some
variables.
The main use of the get\_node\_val function is when the simulation
is aborted for some reason.
Get\_node\_val can then be used to probe the current state of the system.
For example, assuming there is a full\_adder.exe file in the current
working directory, we would get:
\begin{hol}
: let ckt = load_exe "full_adder.exe";
ckt::fsm
: nodes ckt;
..[["cout"],["t4"],["t3"],["t2"],["result"],["cin"],["t1"],["b"],["a"]]
: fanin ckt "cout";
["t2","t3","t4"]
: fanout ckt "t3";
["cout"]
: fanout ckt "t4";
["cout"]
: fanout ckt "a";
["t3","t2","t1"]
: get_node_val ckt "t2";
..(T,T)
\end{hol}

If the user also loads in the "default.fl" standard library, there is
the very useful function {\em excitation} that also takes an fsm model
and a node name, but that returns the next state function for binary inputs.
Note, however, that the current version of the excitation function
only works correctly for unit delay nodes.
\begin{hol}
: load "defaults.fl";
-Loading file defaults.fl
T

: excitation ckt "cout";
....
Trace started for node: cout 
        Current value:X
.Time: 1
.Trace: Node cout  at time 1: t4' + t3' + t2'
Time: 2

Trace ended for node: cout 
"t4' + t3' + t2'"
\end{hol}
We will return to this in Section~\ref{vossrc}.

\subsection{Symbolic Trajectory Evaluation}

There is actually only one built-in command for symbolic trajectory
evaluation called FSM.
In general, FSM determines, through symbolic trajectory evaluation, whether
an antecedent/consequent pair hold in for some circuit.
FSM will return a Boolean function that gives the condition for the
verification to succeed.
For most applications the desired return value is T.

In general FSM is invoked as
\begin{hol}
FSM options fsm ant_list cons_list trace_list
\end{hol}
where options is a string that can contain a combinations of the
following flags:
\begin{description}
\item[-a] Abort the verification at the first antecedent or consequent failure.
If the verification is aborted, FSM will return a Boolean function that
gives {\em the condition for this failure to manifest itself}.
Note that this is contrary to FSM's usual behavior which is to return
the Boolean function that gives the conditions for the verification to
succeed.
\item [-m n]
Abort the verification after reaching time n.
\item [-i]
Allow antecedent failures.
In other words, compute a straight implication.
The normal behavior of the verification process is to disallow
antecedent failures.
Thus the default verification condition is both to check that every
trajectory the circuit can go thorough that is consistent with the
antecedent is also consistent with the consequent, and that there
is at least one (real) circuit trajectory that is consistent
with the antecedent.
\item[-w]
Do not print out warning messages.
\item[-t s]
In addition to printing out trace messages on stderr, also send
the trace events in Postscript format to the file s.
By previewing or printing out the file the user gets a waveform diagram for
the traced signals.
\item[-T s]
Same as -t, but generate Postscript code in landscape mode.
\end{description}

The second argument to FSM must be an object of the fsm type representing
a circuit that is to be simulated.

The ant\_list and cons\_list are  both lists of five-tuples.
Each five-tuple is of the form $(g,n,v,s,t)$, where $g$ is a Boolean function
denoting the domain for which this assertion/check should be carried out,
$n$ is the name of a node, $v$ is the value to be asserted/checked, and
$s$ and $t$ denote the start and stop times for this assertion/check
respectively.

Finally, the last argument to FSM is a list of triples.
Each triple is of the form $(n,s,t)$, where $n$ is a name of a node
to be traced and $s$ and $t$ are the start and stop times for this
trace respectively.

Of course, in practice, it would be quite tedious to have to
write all specifications in terms of lists of five-tuples.
Consequently, a small language (actually a small set of useful functions)
has been defined in the library file "verification.fl".
These functions make it much easier to write specification.
However, it should be remembered that when the verification
is actually performed, all these higher level constructs gets translated
down to the two lists of five-tuples.

For more details on how to use the FSM function, we refer the reader to the
tutorial section.

