\documentstyle[11pt]{article}

\makeindex

\textwidth 165.1truemm
\textheight 232truemm
\oddsidemargin  0truemm
\evensidemargin 0truemm
\topmargin 0truemm
\headheight 0truemm
\headsep 0truemm

\title{Voss --- A Formal Hardware Verification System\\
User's Guide\\
\mbox{}
% \\
%{\large Technical Report 93-45}
}

\author{
\parbox[t]{3.0in}{%
\begin{center}
Carl-Johan H. Seger\\
Department of Computer Science\\
University of British Columbia\\
Vancouver, B.C. V6T 1Z4 Canada \\
Email: seger@cs.ubc.ca
\end{center}}}

\input hol_macros
\input defs


% ====================================================================
%		Start of paper
% ====================================================================


\begin{document}
%\special{!userdict begin /bop-hook{gsave 250 120 translate 65 rotate
%/Times-Roman findfont 160 scalefont setfont
%0 100 moveto 0.7 setgray (DRAFT) show
%/Times-Roman findfont 72 scalefont setfont
%10 0 moveto 0.7 setgray (Do not distribute) show grestore}def end}

\maketitle{}
\begin{abstract}

The Voss system is a formal verification system aimed primarily at
hardware verification.
The special emphasis of the system is to support multi-level
verification.
At the lower levels, symbolic trajectory evaluation (STE)
is used as the main verification engine.
At higher levels, the results obtained by STE is then used for
more general model checking and theorem proving.
The system supports effectively a range of verification approaches
ranging from completely informal simulation based techniques to
sophisticated multi-level verification approaches.
 
The Voss system consists of a collection of programs.
The main program, and the heart of the system, is called fl.
Fl consists essentially of three main parts: a general, strongly typed,
functional language, an efficient implementation of Ordered Binary Decision
Diagrams (OBDDs) built into the functional language, and a symbolic
simulation engine with comprehensive delay modeling capabilities.
Since the interface language to fl is a fully general
functional language in which Ordered Binary Decision Diagrams (OBDDs) has
been built in, the verification system is not only
useful for carrying out STE, but also for experimenting with
various verification (formal and informal) techniques that require
the use of OBDDs.

This document is intended as both a user's guide and (to some extent)
a reference guide.
It describes the alpha release 1.8 of the Voss system.
\end{abstract}


\newpage
\tableofcontents   % Uncomment for final version to generate table of content
\newpage
\listoffigures     % Uncomment for final version to generate list of figures
\listoftables      % Uncomment for final version to generate list of tables
\newpage
\input 01_preface
\newpage
\part{Introduction by Examples}
\newpage
\input 02_background
\newpage
\input 03_fl
\newpage
\input 04_tutorial
\newpage
\input 05_vossprover
\newpage
\part{Reference Manual}
\newpage
\input 06_fl-syntax
\input 07_built-in-fl
\newpage
\input 08_vossrc
\newpage
\input 09_standard-library
\newpage
\appendix
\input 10_AMD2901spec
\newpage
\input 11_switch_level
\newpage
\input 12_sim_format
\newpage
\input 13_ntk_format
\newpage
\input 14_sil_format
\newpage
\input 15_vhdl
\newpage
\input 16_biblio
\newpage
\input glossary
\end{document}
